EEG Waveform Analysis with applications to Brain Computer Interfaces

### 2018-08-13 sensors.tex
Added some equations.

@@

Pursuit algorithms refer, in their many variants, as blind source separation(REF) techniques that assume that the EEG signal is a linear combination of different sources that comes from template's dictionaries.  The underlying idea is to find the best template out of a dictionary that matches against the signal.  The algorithm works iteratively, starting from the raw signal and then, the template is substracted from the signal in Graham-Schmidt approach and the process iterates. (ref Cohen book, Sanei book, Mallat and Zhang 1993).

Two variants of this algorithm are evaluated. In \textit{MP 1} the dictionary is directly constructed with the normalized templates from Fig \ref{fig:dictionaryfig}.  In \textit{MP 2} the coefficients of  Daubechies least-asymetric wavelet with 2 vanishing moments atoms are used to construct the dictionary.  To construct each feature a signal is decomposed in its coefficients and then they can be compared by any distance function.

@@

Bond and Pompe Permutation Entropy has been extensively used in EEG processing, particularly for Epilepsy pre-ictal detection.  This method generates a code based on the orderly arrangement of sequential samples, and then derives a metric which is based on the amount of entropy of each code within the signal.   For instance, a pure random signal, will achieve the maximum entropy value due to the probability of each code being equal for all of them. This method is the best representative of Waveform Complexity: Lempel-Ziv method L-Z complexity are other variants which use a different definition (Permutation Entropy a new feature for brain computer interfaces).

@@

@@

>>>>

\begin{table}[H]
\caption{Subjective EEG characterization methods found in the surveyed literature.}
\centering
%% \tablesize{} %% You can specify the fontsize here, e.g.  \tablesize{\footnotesize}. If commented out \small will be used.
\begin{tabular}{ccc}
\toprule
\textbf{Method}	& \textbf{Phenomena} & \textbf{Reference}	\\
\midrule
SVM     & m-c  & $50\%$ \\
NN       & m-c  & $45\%$ \\
HGO  & PO8 & $25\%$ \\
MP 1 & PO8 & $25\%$ \\
PE     & Cz & $20\%$ \\
RS     & Fz & $15\%$ \\
RS     & m-c  & $15\%$ \\
MP 2 & PO7 & $10\%$ \\
SHCC & P4 & $10\%$ \\
NN s-c & Fz & $10\%$ \\
\bottomrule
\end{tabular}
\label{tab:results}
\end{table}

@@

\textit{Pursuit} algorithms refer, in their many variants, as blind source separation \citep{Comon2010} techniques that assume that the EEG signal is a linear combination of different sparse sources extracted from a template's dictionaries.  \textit{MP}, the most representative of this algorithms, is a greedy variant that decomposes a signal into a linear combination of waveforms, called atoms, that are well localized in time and frequency \citep{ChandranKS2016}.  Given a signal, this optimization technique, tries to find the indexes of atoms and their weights (contributions) that minimize,

\begin{equation}
\varepsilon =  \left\lVert   x(n,c) - \sum_{i=1}^{m} w_i g_{yi}   \right\rVert
\label{eq:mperror}
\end{equation}

which is the error from the signal approximation constructed by the weighted atoms.  The algorithm goes by first setting the approximation signal by the original signal itself,

\begin{equation}
\tilde{x_{0}}(n,c) = x(n,c), k \leftarrow 1
\label{eq:multiclassificationrow}
\end{equation}

Hence, it finds the best template, out of the dictionary, that matches current approximation (original signal).

\begin{equation}
g_{yk}(n) = \arg \max_{g_{yi}} \left\lVert < \tilde{x_{k}}(n,c), g_{yi} > \right\rVert 
\label{eq:multiclassificationrow}
\end{equation}

where $g_{yi}$ are all the available scaled, translated and modulated atoms from the dictionary.  This determines the atom selection process, and their contribution is calculating based on

\begin{equation}
w_{k} =  \frac{< \tilde{x_{k}}(n,c), g_{yk} >}{g_{yk}}
\label{eq:multiclassificationrow}
\end{equation}

with $k$ representing the index of the selected atom $g_{yk}$.  Finally the contribution of each atom is subtracted from the next approximation, refining the approximation process \citep{Cohen2008,Sanei1997, Mallat1993}

\begin{equation}
\tilde{x_{k+1}}(n,c)=  \tilde{x_k}(n,c) - w_{k} g_{yk} (n), k \leftarrow k + 1
\label{eq:multiclassificationrow}
\end{equation}

Two variants of this algorithm are evaluated. In \textit{MP 1} the dictionary is directly constructed with the normalized templates from Fig \ref{fig:dictionaryfig}.  In \textit{MP 2} the coefficients of  Daubechies least-asymetric wavelet with 2 vanishing moments atoms are used to construct the dictionary.  The first version, the matching against the template is evaluated according to Equation \ref{eq:mperror} directly, whereas for the latter to craft each feature is crafted by decomposing the signal in its coefficients and building a vector with them.

\begin{equation}
f =  {\bigg ( w_{i} \bigg )}_{1}^{m} 
\label{eq:multiclassificationrow}
\end{equation}

@@

Bond and Pompe Permutation Entropy has been extensively used in EEG processing, particularly for Epilepsy pre-ictal detection \citep{Bandt2002}.  This method generates a code based on the orderly arrangement of sequential samples, and then derives a metric which is based on the amount of entropy of each code within the signal.   For instance, a pure random signal, will achieve the maximum entropy value due to the probability of each code being equal for all of them. This method is the best representative of Waveform Complexity: Lempel-Ziv method L-Z complexity are other variants which use a different definition (Permutation Entropy a new feature for brain computer interfaces).

A signal represented by

\begin{equation}
(x_1,x_2,...,x_{N})
\label{eq:pesignal}
\end{equation}

is resampled by $\tau$ values doing

\begin{equation}
(x_n,x_{n+\tau},x_{n+2 \tau}...,x_{n+(N-1)\tau})
\label{eq:pesignal}
\end{equation}

@@

A signal, can be represented by a list of ordered-pairs $e$,

\begin{equation}
e = \left[ (x,y)_{1}, (x,y)_{2}, ..., (x,y)_{N} \right]
\label{eq:shccdelta}
\end{equation}

\noindent and it can be divided into $G$ different blocks.  These blocks are obtained by resampling the original signal from the index

\begin{equation}
g = \lfloor n + ( m \Delta ) + 0.5 \rfloor
\label{eq:shccdelta}
\end{equation}

\noindent with $n$ being the original sampling index, and $ 1 \leq t \leq N $.  On the other hand, $\Delta$ can be obtained by

\begin{equation}
\Delta = \bigg \lceil \frac{N}{G+1} \bigg \rceil
\label{eq:shccdelta}
\end{equation}

\noindent with $ G < N $. The value $m$ can be derived from

\begin{equation}
m = sign \bigg (  \frac{N-1}{\Delta} \bigg )  \bigg \lfloor \left\lvert \frac{T-1}{\Delta} \right\lvert \bigg \rfloor
\label{eq:shccdelta}
\end{equation}

This resampling produces a new sequence of values,

\begin{equation}
e = \left[ (x',y')_{1}, (x',y')_{s}, ..., (x',y')_{G} \right]
\label{eq:shccdelta}
\end{equation}

Next step is the normalization of each paired-values according to

\begin{equation}
\textbf{x} = \frac{1}{\max(x') - min(x')} ( \textbf{x'} - \min(x') \textbf{1} )
\label{eq:shccdelta}
\end{equation}

\begin{equation}
\textbf{y} = \frac{1}{\max(y') - min(y')} ( \textbf{y'} - \min(y') \textbf{1} )
\label{eq:shccdelta}
\end{equation}

\noindent effectively yielding $x,y \in [0,1]$.

Finally, the feature is constructed by calculating the point-to-point slope against the horizontal plane,

\begin{equation}
f = \bigg (  \frac{y_{s}-y_{s-1}}{x_{s}-x_{s-1}}  \bigg )_{2}^{G} 
\label{eq:shccdelta}
\end{equation}

@@

\begin{equation}
f = {\bigg ( h(i,j,\theta) \bigg )}_{i,j,\theta}
\label{eq:multiclassificationrow}
\end{equation}
